---
title: "Hold Out Method vs. Cross Vaildation"
author: "Tadros Salama"
date: "2/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(AmesHousing) 
library(caret)
ames <- make_ames()
```

## This analysis uses 3 linear models performances on predicting sales price of a home to compare different testing and validating processes. The first method will be the hold out method approach for `lm1, lm2, lm3`, and the second, a 10-fold cross validation using `train()` from the R package (caret)[https://cran.r-project.org/web/packages/caret/caret.pdf] on `cv_lm1, cv_lm2, cv_lm3`.

Data from R package (AmesHousing)[https://github.com/topepo/AmesHousing]

### Spliting data into 70% train and 30% test
```{r}
set.seed(123)
train_index <- sample(1:nrow(ames), round(nrow(ames) * 0.7))
ames_train <- ames[train_index, ]
ames_test <- ames[-train_index, ]
```

### Model 1, single predictor variable - total rooms in a house
### Model 2, two predictor variables - total rooms & year the house was built
### Model 3, three perdictor variables- total rooms above ground, year built, and overall condition of house
```{r}
lm1 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, ames_train)
lm2 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built + TotRms_AbvGrd, ames_train)
lm3 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built + TotRms_AbvGrd + Overall_Cond, ames_train)
```


```{r}
pred_lm1 <- predict(lm1, newdata = ames_test)
pred_lm2 <- predict(lm2, newdata = ames_test)
pred_lm3 <- predict(lm3, newdata = ames_test)

RMSE <- function(y, y_hat) {
  sqrt(mean((y - y_hat)^2))
}

RMSE(ames_test$Sale_Price, pred_lm1)
RMSE(ames_test$Sale_Price, pred_lm2)
RMSE_lm3 <- RMSE(ames_test$Sale_Price, pred_lm3)

```

```{r}
MAE <- function(y, y_hat) {
  mean(abs(y - y_hat))
}

MAE(ames_test$Sale_Price, pred_lm1)
MAE(ames_test$Sale_Price, pred_lm2)
MAE(ames_test$Sale_Price, pred_lm3)
```

```{r}
df_pred_lm3 <- data.frame(
  actual = ames_test$Sale_Price / 1000,
  predLm3 = pred_lm3 / 1000
  )

ggplot(df_pred_lm3, aes(actual, predLm3)) +
  geom_point(shape =1) +
  geom_abline(intercept = 0, slope = 1, color = 'blue') +
  xlab("Actual Sale Price (in thousands)") +
  ylab("Predicted Sale Price (in thousands)") +
  ggtitle("Lm3") 
  
```

## Using Cross-Validiation to compare model perfromance
### Model 4, single predictor variable - total rooms in a house
### Model 5, two predictor variables - total rooms & year the house was built
### Model 6, three perdictor variables- total rooms above ground, year built, and overall condition of house
```{r}

set.seed(123)
cv_lm1 <- train(
  Sale_Price ~ TotRms_AbvGrd,
  data = ames,
  method = 'lm',
  trControl = trainControl(method = 'cv', number = 10)
)
cv_lm1
```

```{r}
set.seed(123)
cv_lm2 <- train(
  Sale_Price ~ TotRms_AbvGrd + Year_Built,
  data = ames,
  method = 'lm',
  trControl = trainControl(method = 'cv', number = 10)
)
cv_lm2
```

```{r}
set.seed(123)
cv_lm3 <- train(
  Sale_Price ~ TotRms_AbvGrd + Year_Built + Overall_Cond,
  data = ames,
  method = 'lm',
  trControl = trainControl(method = 'cv', number = 10)
)
cv_lm3
```








































