---
title: 'Bitcoin Price Forecast'
author: "Tadros Salama"
date: "4/28/2021"
output:
  rmdformats::downcute:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: false
    gallery: false
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp3)
library(faraway)
library(lubridate)
library(gridExtra)
btc <- read_csv("~/Desktop/474 project/btc.csv")
```

Over the past year the price of bitcoin has risen more than 300%, growth rarely
seen in traditional financial markets. Bitcoin is notorious for it's 
unpredictable price movements and volatility. So what could be more fun than
forecasting the potential price of 1 bitcoin a year out.

There is close to 19 million bitcoin in circulation
with a cap of 21 million bitcoin that can be minted. It's scarcity, similar to 
gold, has been one of it's most appealing qualities too investors.
So in this forecast the price of bitcoin, `PriceUSD` and it's supply, `SplyCur`
will be the two variables I use.

In this forecast I used historical data from [coinmetrics](https://coinmetrics.io/)
tracking daily data from June 2010 till April 2021.


## EDA

```{r include=FALSE}
data_btc <- btc %>%
  select(date, PriceUSD, SplyCur, NVTAdj) %>% filter(date > '2010-07-18')
data_btc <- as_tsibble(data_btc)

btc_log <- data_btc %>% mutate(logPrice = abs(log(PriceUSD)))
head(btc_log)
tail(btc_log)
```

```{r}

p1 <- ggplot(btc_log,aes(SplyCur, logPrice)) + geom_line() + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=F) + coord_trans(x = 'log2')
p2 <- ggplot(btc_log,aes(date, logPrice)) + geom_line() + geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=F) + coord_trans(x = 'log2')
p3 <- ggplot(btc_log,aes(date, SplyCur/1000000)) + geom_line() + ylab('Supply in Circulation')
grid.arrange(p1, p2, p3, ncol=1, nrow =3)
```
### Transformatitons

```{r}
#box_cox transofrmation
lambda <- data_btc %>% 
  select(PriceUSD) %>% 
  features(PriceUSD, features = guerrero) %>%
  pull(lambda_guerrero)
lambda

data_btc %>%  
  select(PriceUSD) %>%
  autoplot(box_cox(PriceUSD, lambda))

#Log transformation
data_btc %>%
  autoplot(log(PriceUSD))

```


## Linear regression model

```{r}
#Overfitting for fun
lm_fit_btcLog <- btc_log %>%
  model(
    lm1 =TSLM(logPrice ~ trend() + season('4years'))
   
    )
lm1 <- btc_log %>%
  model(TSLM(logPrice ~ SplyCur))


augment(lm_fit_btcLog) %>% 
  ggplot(aes(x=date)) + geom_line(aes(y=logPrice, colour = 'data')) +
  geom_line(aes(y=.fitted, colour='fitted'))

lm_fit_btcLog %>%
  forecast(h=365) %>% autoplot(btc_log)

```


## ARIMA

```{r}
fit_arima <- btc_log %>%
  model(ARIMA(logPrice))
accuracy(fit_arima)
fit_arima %>% forecast(h='12 months') %>% autoplot(btc_log)

fit_arima %>% gg_tsresiduals()
glance(fit_arima)

```


## ETS

```{r}
fit_ets <- btc_log %>%
  model(
    ses = ETS(logPrice ~ error("A") + trend("N") + season("N")),
    holt = ETS(logPrice ~ error("A") + trend("A") + season("N")),
    damped = ETS(logPrice ~ error("A") + trend("Ad") + season("N")),
    ets1 = ETS(logPrice)
  ) 

fit_ets <- btc_log %>%
  model(ETS(logPrice))

fit_ets %>%
  forecast(h = '12 months') %>%
  autoplot(btc_log)
fit_ets %>%
  gg_tsresiduals()
```


## Nueral Net

```{r}

nns <- btc_log %>%
  model(NNETAR(logPrice))
nns %>% forecast(h='12 months') %>%
  autoplot(btc_log)
```






